### 2024-01-19 scomatic/reports/scomatic.Rmd
restructure infercnv and scomatic runs into separate subdirectories

---
title: "Mutation loads in rheumatoid arthritis data from SComatic"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: true
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
    do_rerun: TRUE
    celltype_counts: NULL
    mappings: NULL
---

```{css style settings, echo = FALSE}
blockquote {
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 14px;
    border-left: 5px solid #eee;
}
```

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# params <- list(celltype_counts = 'data/scomatic/celltype_counts.tsv', do_rerun = T)
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE,
  cache.path = 'reports/mutation_load_cache/')

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)
```

SComatic was run on scRNA-seq rheumatoid arthritis data using a modified version of kp9 and jc48’s [Nextflow implementation](https://github.com/Teichlab/scripts/tree/bd17ed22ec9e669056862d28c5f12a7e5033e68c/genotyping/scomatic/nextflow). 
Their implementation was intended to capture all germline and somatic variants in order to reconstruct tissue phylogeny, using both shared and private mutations for lineage tracing. In order to achieve this, they set `max_cell_types` to 1e6. I set it to 1. I also forced the pipeline to publish the celltype-split BAMs for each patient, which will be used to assess coverage of the driver genes in each cell type. The pipeline can be used with both GEX and ATAC modalities, hence the `*-GEX` suffix on the output directories.

Here is an example of the output files for donor BRI-401:

```
out/scomatic/BRI_401-GEX
├── BRI_401.calling.step2.intersect.tsv
├── BRI_401.calling.step2.pass.tsv
├── BRI_401.calling.step2.tsv
├── BRI_401.coverage_cell_count.per_chromosome.report.tsv
├── BRI_401.coverage_cell_count.report.tsv
├── cell_callable_sites
│   ├── BRI_401.B_cell_plasma_cell.SitesPerCell.tsv
│   ├── BRI_401.Endothelial_cell.SitesPerCell.tsv
│   ├── BRI_401.Myeloid_cell.SitesPerCell.tsv
│   ├── BRI_401.NK.SitesPerCell.tsv
│   ├── BRI_401.Stromal_cell.SitesPerCell.tsv
│   └── BRI_401.T_cell.SitesPerCell.tsv
└── celltype_bams
    ├── BRI_401.B_cell_plasma_cell.bam
    ├── BRI_401.B_cell_plasma_cell.bam.bai
    ├── BRI_401.Endothelial_cell.bam
    ├── BRI_401.Endothelial_cell.bam.bai
    ├── BRI_401.Myeloid_cell.bam
    ├── BRI_401.Myeloid_cell.bam.bai
    ├── BRI_401.NK.bam
    ├── BRI_401.NK.bam.bai
    ├── BRI_401.Stromal_cell.bam
    ├── BRI_401.Stromal_cell.bam.bai
    ├── BRI_401.T_cell.bam
    └── BRI_401.T_cell.bam.bai
```

### Sample metadata

```{r md}
ct_counts <-
    params$celltype_counts %>%
    readr::read_tsv()
mappings <-
    params$mappings %>%
    readr::read_csv()
```

### Celltype counts

```{r plot_cts_vs_ids, results = F, message = F}
p_dat <-
    ct_counts %>%
    dplyr::group_by(celltype) %>%
    dplyr::mutate(
        n_cells_per_celltype = sum(n_cells_per_id_per_celltype)) %>%
    dplyr::group_by(id) %>%
    dplyr::mutate(
        n_cells_per_id = sum(n_cells_per_id_per_celltype)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
        id = forcats::fct_reorder(id, n_cells_per_id),
        celltype = forcats::fct_reorder(celltype, -n_cells_per_celltype),
        dplyr::across(
          .cols = dplyr::starts_with('n_'), 
          .fns = ~ prettyNum(.x, big.mark = ",", scientific = FALSE),
          .names = '{.col}_label'))

p_ids_x_cts <-
  p_dat %>%
  ggplot(aes(x = id, y = celltype, 
             colour = celltype)) +
  geom_point(aes(size = n_cells_per_id_per_celltype, 
                 alpha = n_cells_per_id_per_celltype)) +
  scale_alpha(range = c(0.5, 1)) +
  scale_y_discrete(position = 'right')  +
  scale_x_discrete(position = 'bottom', guide = guide_axis(angle = -90)) +
  scale_size_continuous(range = c(1, 8)) +
  theme_linedraw() +
  theme(legend.position = 'left', legend.title = element_blank(),
        axis.text.x = ggtext::element_markdown(angle = 90, vjust = 0.5),
        axis.title = element_blank(),
        strip.placement = 'outside') 
```

### PASS/FAIL criteria

```{r pass_crit}
prefilter <-
    mappings$id %>%
    purrr::set_names() %>% 
    purrr::map(function(id) {
        step2 <- 
            paste0(params$output_dir, id, '-GEX') %>%
            list.files(pattern = '.calling.step2.tsv', full.names = T) 
        if (length(step2)) { readr::read_tsv(step2, comment = '##') }
    }) %>%
    dplyr::bind_rows(.id = 'id')
prefilter %>%
    dplyr::mutate(FILTER = FILTER == 'PASS') %>%
    dplyr::group_by(id) %>%
    dplyr::mutate(
        n_muts = dplyr::n(),
        id = forcats::fct_reorder(id, -n_muts)) %>%
    ggplot(aes(x = id, fill = FILTER)) +
    geom_bar() +
    theme(axis.text.x = element_text(angle = -90))
```

### Load called mutations

```{r load_som_muts}
scom <- xfun::cache_rds({

  # initiate lists
  scom <- list(
    all_muts = list(),
    som_muts = list(),
    coverages = list(),
    sc_muts = list(),
    annovar = list()
  )

  # load somatic mutations and coverages for each ID
  dir <- params$output_dir
  list.files(dir, pattern = '-GEX$') %>%
    purrr::walk(function(subdir) {

      id <- gsub('-GEX', '', subdir)
      file_prefix <- paste0(dir, '/', subdir, '/', id)
      message(id)

      # get all called muts
      scom$all_muts[[id]] <<-
        paste0(file_prefix, '.calling.step2.pass.tsv') %>%
        readr::read_tsv(comment = '##') %>%
        # separate longer the Cell_types column and pivot longer the
        # calling columns and then separate wider the INFO 
        # -> produces one row per variant per celltype
        tidyr::pivot_longer(!c(`#CHROM`:INFO), names_to = 'ct', values_to = 'call') %>%
        tidyr::separate_rows(Cell_types, VAF, sep = ',') %>%
        tidyr::separate(
          'call', 
          into = paste0('ct_', c('DP', 'NC', 'CC', 'BC', 'BQ', 'BCf', 'BCr')),
          sep = '\\|',  remove = F, convert = T) %>%
        readr::type_convert() %>%
        dplyr::filter(ct == Cell_types) %>%
        # get celltype names
        dplyr::mutate(
          annotation_level1_alias = gsub(col_suffix, '', Cell_types))

      # get called somatic mutations (exclusive to celltypes)
      scom$som_muts[[id]] <<-
        scom$all_muts[[id]] %>%
        # get vars shared by <= max_n_celltypes
        dplyr::group_by(`#CHROM`, Start, End, REF, ALT) %>%
        dplyr::filter(dplyr::n_distinct(annotation_level1_alias) == 1) %>%
        # get median VAF per celltype
        dplyr::group_by(annotation_level1_alias) %>%
        dplyr::mutate(median_VAF_per_id_per_celltype = median(VAF)) %>%
        dplyr::ungroup()

      # get callable sites per celltype
      scom$coverages[[id]] <<-
        paste0(file_prefix, '.coverage_cell_count.per_chromosome.report.tsv') %>%
        readr::read_tsv() %>%
        # get celltype names
        dplyr::mutate(
          annotation_level1_alias = gsub(col_suffix, '', Cell_types))

      # get annovar gene annotations
      scom$annovar[[id]] <<-
        paste0(dir, '/', subdir, '/annovar/', 
               id, '.variants.annovar.hg38_multianno.csv') %>%
        readr::read_csv()

      # get somatic mutations at single cell resolution
      scom$sc_muts[[id]] <<-
        paste0(dir, '/', subdir, '-genotypes') %>%
        list.files(full.names = T) %>%
        purrr::map(function(file) {
          # message(file)
          sc_df <-
            readr::read_tsv(file, ) %>%
            dplyr::filter(Cell_type_expected == Cell_type_observed,
                          ALT_expected == Base_observed) %>%
            # get celltype names
            dplyr::mutate(
              annotation_level1_alias = gsub(
                paste0('_', id), '', Cell_type_observed))

          # get callable sites per cell
          ct <- unique(sc_df$Cell_type_observed)
          if (length(ct) != 1) {
            message(paste(length(ct), 'celltypes in the single cell genotypes file! Should be 1!'))
            message(paste('File', file, 'will be skipped!'))
            sc_df <- tibble::tibble()
          } else {
            sc_coverage <- 
              paste0(dir, '/', subdir, '/cell_callable_sites/', 
                     id, '.', ct, '.SitesPerCell.tsv') %>%
              readr::read_csv()
            # add single cell-level coverage
            sc_df <- dplyr::left_join(sc_df, sc_coverage, by = 'CB')
          }

          # return
          sc_df

        }) %>%
        dplyr::bind_rows()

    })

  # add sample and celltype metadata
  scom <-
    scom %>% 
    purrr::map(function(df_ls) {
      df_ls[sapply(df_ls, function(x) dim(x)[1]) > 0] %>%
        dplyr::bind_rows(.id = 'id')
    }) %>%
        # # add celltype info
        dplyr::left_join(ct_info) %>%
        # # add ID info
        dplyr::left_join(sm) %>%
        # add grouping structure
        dplyr::group_by(# sample info
                        id, donor_id, sex, age_bracket, age, X10X_version, n_cells_per_id,
                        # celltype info
                        annotation_level1_alias, annotation_level1, n_cells_per_id_per_celltype)
    })

  # return
  scom

}, file = 'scom.rds', rerun = params$do_rerun)

# assign dfs globally
all_muts <- xfun::cache_rds({scom$all_muts}, file = 'all_muts.rds', rerun = do_rerun)
som_muts <- xfun::cache_rds({scom$som_muts}, file = 'som_muts.rds', rerun = do_rerun)
coverages <- xfun::cache_rds({scom$coverages}, file = 'coverages.rds', rerun = do_rerun)
sc_muts <- xfun::cache_rds({scom$sc_muts}, file = 'sc_muts.rds', rerun = do_rerun)
annovar <- xfun::cache_rds({scom$annovar}, file = 'annovar.rds', rerun = do_rerun)
```