### 2024-01-30 scomatic/reports/driver_coverage.Rmd
add driver preprocessing script to find exons, will eventually add to driver_coverage.Rmd, fix infercnv output pattern matching

---
title: "Mutation loads in rheumatoid arthritis data from SComatic"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: true
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
    rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') %>% paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/') ; setwd(wd)
# rmarkdown::render('reports/driver_coverage.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = T))

knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE,
  cache.path = 'reports/driver_coverage_cache/')

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)

# dev.off() if there is a device
dev_off_if <- function() {if (length(dev.list()) != 0) { dev.off() }}
```

### Drivers

Genes of interest in rheumatoid arthritis (hereafter referred to as 'drivers') are defined as genes that are found to undergo significant selection (q-value < 0.2) according to the whole exome and targeted NanoSeq data. The coverage of these sites in the dataset, across samples and celltypes, will be checked.

First, we read in the dNdS results and filter by q-value.

```{r read_dnds, warning = F, message = F}
# read in dnds results
muts <-
  'data/dnds_nanoseq/' %>%
  list.files(pattern = 'nanoseq') %>%
  purrr::map(function(assay) {
    list.files(paste0('data/dnds_nanoseq/', assay), 
               pattern = '_dnds_sel_cv.tsv', 
               full.names = T) %>%
      purrr::map(function(file) {
        file %>%
          readr::read_tsv(show_col_types = F) %>%
          dplyr::mutate(
            assay = assay,
            celltype = file %>% 
              basename() %>% 
              gsub('_dnds_sel_cv.tsv', '', .))
      }) %>%
      dplyr::bind_rows()
  }) %>%
  dplyr::bind_rows() %>%
  dplyr::select(assay, celltype, everything())

# get genes with strong evidence (q-value)
drivers <-
  muts %>%
  dplyr::filter(qglobal_cv < 0.2) %>%
  dplyr::rename(gene = gene_name)

# manually write aliases for genes that are not matched
# HIST1H3B -> H3C2
# HIST1H1E -> H1-4
# CRIPAK   -> This record has been withdrawn by NCBI, after discussions with 
#             CCDS collaborators. It was decided that this locus is not an 
#             independent gene.
drivers <-
  drivers %>%
  dplyr::mutate(
    gene_fixed = dplyr::case_when(
      gene == 'HIST1H3B' ~ 'H3C2',
      gene == 'HIST1H1E' ~ 'H1-4',
      TRUE ~ gene
    )
  )
```

```{r plot_drivers, class.source = "fold-hide", warning = F, message = F}
drivers %>%
  dplyr::distinct(celltype, gene, assay, qglobal_cv) %>%
  dplyr::mutate(neglog10q = -log10(qglobal_cv)) %>%
  ggplot(aes(x = tidytext::reorder_within(gene, -neglog10q, celltype), 
             y = neglog10q, fill = celltype)) +
  geom_col(position = 'dodge') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0)) +
  facet_grid(assay ~ celltype, scales = 'free_x', space = 'free_x') +
  tidytext::scale_x_reordered() +
  labs(x = 'gene', y = '-log10(p)')
```

```{r drivers_df, echo = F}
drivers %>%
  knitr::kable("html") %>% 
  kableExtra::kable_styling("striped", full_width = F) %>% 
  kableExtra::scroll_box(width = "100%", height = "300px")
```

In order to calculate their coverage in the BAMs, we must get their positions.

```{r get_driver_pos}
# load biomart
library(biomaRt)

# load ref
ensembl <- useEnsembl(biomart = 'ensembl',
                      dataset = 'hsapiens_gene_ensembl')

# get positions
drivers_pos <-
  getBM(attributes = c('hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'),
        filters = 'hgnc_symbol', values = unique(drivers$gene),
        mart = ensembl) %>%
  # fix colnames
  dplyr::mutate(gene = hgnc_symbol, 
                chr = chromosome_name, 
                start = start_position,
                end = end_position) %>%
  # remove weird chromosomes
  dplyr::filter(chr %in% c(as.character(1:22, 'X', 'Y')))

# save positions
drivers_pos %>% 
  dplyr::group_by(gene) %>%
  dplyr::summarise(coords = 
                     paste0(unique(chromosome_name), ':', 
                            min(start_position), '-', 
                            max(end_position))) %>%
  readr::write_tsv('data/driver_genes/driver_gene_coords_for_coverage.tsv')
```

### Calculate coverage, by sample and by sample-x-celltype

We use `samtools depth` to get the coverage of each site in the drivers, both across each sample and within each cell type in each sample.

```{bash cov, eval = T}
# activate conda env
mamba activate jupy

# dirs
input_dir=/lustre/scratch126/casm/team268im/mp34/scRNAseq_data/RA_ZhangEtal2023/cellranger_output
ct_input_dir=out/Zhang2023/default_thresholds/
output_dir=out/Zhang2023/coverage/
drivers=data/driver_genes/driver_gene_coords_for_coverage.tsv
mkdir -p $output_dir

# iterate over each bam file and pileup specific regions
(
  echo -e 'chr\tpos\tdepth\tgene\tsample' ;
  for bam in $(find $input_dir -iname 'possorted_genome_bam.bam'); do
  	sample=$(echo $bam | cut -d "/" -f 10 | sed 's/\-/_/g')
    while read -r gene coords ; do
      samtools depth -a -r $coords -Q 30 $bam |
      awk -v sample=$sample -v gene=$gene '{print $0"\t"gene"\t"sample}'
    done < <(sed 1d $drivers) ;
  done
) | cat > $output_dir/drivers_samtools_depth.tsv

# iterate over celltype-specific bam files and pileup specific regions
(
  echo -e 'chr\tpos\tdepth\tgene\tsample\tcelltype' ;
  for sample_dir in $ct_input_dir/*-GEX ; do
    sample=$(basename $sample_dir | sed 's/\-GEX$//g')
    for ct_bam in $sample_dir/celltype_bams/*.bam ; do
      ct=$(basename $ct_bam | cut -f2 -d.)
      while read -r gene coords ; do
        samtools depth -a -r $coords -Q 30 $ct_bam |
        awk -v sample=$sample -v gene=$gene -v ct=$ct \
        '{print $0"\t"gene"\t"sample"\t"ct}'
      done < <(sed 1d $drivers) ;
    done
  done
) | cat > $output_dir/drivers_samtools_depth_by_celltype.tsv
```

We read in the results at both levels.

```{r read_cov}
cov <- xfun::cache_rds({
  cov_files <-
    list.files(
      'out/Zhang2023/coverage/', 
      pattern = 'drivers_samtools_depth.*.tsv',
      full.names = T) 
  cov <-
    cov_files %>%
    setNames(cov_files %>% basename() %>% tools::file_path_sans_ext()) %>%
    purrr::map(readr::read_tsv) %>%
    dplyr::bind_rows(.id = 'lvl') %>%
    dplyr::mutate(chr = as.character(chr))
  cov
}, file = 'cov.rds', rerun = params$rerun)
```

### Annotate exons and introns

We use the reference object from DNDSCV to delimit intronic and exonic regions of the genes.

```{r features}
# annotate gene features
cov <- xfun::cache_rds({

  # load rda to get ref_cds object from DNDSCV
  refcds_38 <- '/lustre/scratch125/casm/team268im/fa8/119/DNDSCV_COVARIATES/refcds_GRCh38-GencodeV18+Appris.rda'
  load(refcds_38)

  # set gene names
  ref_cds <-
    RefCDS %>%
    setNames(RefCDS %>% purrr::map(~ .x$gene_name) %>% unlist())

  # annotate exons
  exons <-
    ref_cds[c(drivers$gene)] %>%
    purrr::map(~ .x$intervals_cds %>% 
                 tibble::as_tibble() %>%
                 dplyr::transmute(chr = .x$chr, start = V1, end = V2) %>%
                 dplyr::mutate(exon = paste0('exon', dplyr::row_number()))) %>%
    dplyr::bind_rows(.id = 'gene') %>%
    dplyr::mutate(pos = purrr::map2(start, end, ~ .x:.y)) %>%
    tidyr::unnest(cols = 'pos') %>%
    dplyr::select(-start, -end)
  cov <-
    cov %>%
    dplyr::left_join(exons, by = c('chr', 'pos', 'gene')) %>%
    dplyr::mutate(exon = ifelse(is.na(exon), 'intron', exon),
                  feature = ifelse(exon == 'intron', 'intron', 'exon'))

  # return
  cov

}, file = 'cov_w_features.rds', rerun = params$rerun)
```

```{r plots}
# get exon coords for plotting
exon_coords <- xfun::cache_rds({
  exon_coords <-
    cov %>%
    dplyr::filter(feature == 'exon') %>%
    dplyr::group_by(gene, exon) %>%
    dplyr::summarise(start = min(pos), 
                     end = max(pos))
  exon_coords
}, file = 'exon_coords.rds', rerun = params$rerun)

# function: plot all
plot_all <- function(p_df, p_exon_coords, p_title) {
  ggplot() +
    geom_rect(
      data = p_exon_coords,
      aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
      colour = 'lightblue', fill = 'lightblue'
    ) +
    geom_col(
      data = p_df,
      aes(x = pos, y = depth, fill = feature)) +
    theme_classic() +
    theme(panel.border = element_rect(fill = NA),
          panel.spacing = unit(.05, 'lines')) +
    scale_fill_manual(values = c(exon = 'darkblue', intron = 'darkgrey')) +
    labs(title = paste(p_title, '(all)')) +
    facet_grid(celltype ~ ., space = 'free_x', scales = 'free_x') 
}

# function: plot exons 
plot_exons <- function(p_df, p_title) {
  p_df %>%
    # get only exonic sequence
    dplyr::filter(feature == 'exon') %>%
    # fix exon ordering
    dplyr::mutate(
      exon = exon %>% forcats::fct_reorder(exon %>% gsub('exon', '', .) %>% as.numeric())
    ) %>%
    ggplot(aes(x = pos, y = depth, group = celltype)) +
    geom_col(fill = 'darkblue') +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.spacing = unit(.05, 'lines'),
          panel.border = element_rect(fill = NA)) +
    labs(title = paste(p_title, '(exons)')) +
    facet_grid(celltype ~ exon, space = 'free_x', scales = 'free_x') 
}

# create list of plots by gene x sample
p <- xfun::cache_rds({

  # initiate plot list
  p <- list()
  n_cts <- list()

  cov %>% 
    dplyr::group_by(gene, sample) %>%
    dplyr::group_split() %>%
    purrr::map(function(df_i) {

      # for testing:
      # cov %>% dplyr::group_by(gene, sample) %>% dplyr::group_split() %>% {.[[1]]} -> df_i

      # get info
      i <- 
        df_i %>%
        dplyr::distinct(gene, sample)
      info <- t(i[1,]) %>% paste(collapse = '.')
      p[[info]] <<- list()

      # get number of celltypes for plot dims
      p[[info]]$n_cts <<- 
        df_i %>% 
        dplyr::filter(!is.na(celltype)) %>% 
        dplyr::pull(celltype) %>% 
        dplyr::n_distinct()

      # plot by sample and by celltype
      list(by_sample = list(lvl = 'drivers_samtools_depth', title_suffix = ''),
           by_celltype = list(lvl = 'drivers_samtools_depth_by_celltype', title_suffix = '_by_celltype')) %>%
        purrr::map(function(x){

          # plot all
          p[[info]][[paste0('all', x$title_suffix)]] <<-
            plot_all(
              p_df = df_i %>% dplyr::filter(lvl == x$lvl),
              p_exon_coords = exon_coords %>% dplyr::filter(gene == i$gene),
              p_title = paste0(info, x$title_suffix)
            )

          # plot exons only
          p[[info]][[paste0('exons', x$title_suffix)]] <<-
            plot_exons(
              p_df = df_i %>% dplyr::filter(lvl == x$lvl),
              p_title = paste0(info, x$title_suffix)
            )

        })

    })

  # return 
  p

}, file = 'plots.rds', rerun = params$rerun)
```

```{r knit_plots, echo = F, warning = F, message = F, results = 'asis'}
chunks <- c()

for (info in names(p)) {
  chunks <- c(chunks, paste0('\n####', info))
  message(info)

  knit_plot <- c(
    '',
    paste0('```{r ', info, ', class.source = "fold-hide", warning = F, message = F, ',
           'fig.dim = c(20, ', 4 + 4 * p[[info]]$n_cts, ')}'),
    paste0('p_i <- p[["', info, '"]]'),
    paste0('grid.arrange(p_i$all, p_i$all_by_celltype, p_i$exons, p_i$exons_by_celltype, ncol = 1, ',
           'heights = c(2, p_i$n_cts * 2, 2, p_i$n_cts * 2))'),
    '```',
    '')
  knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
  dev_off_if()

  chunks <- c(chunks, knitted_plot)
}
dev_off_if()

# evaluate chunks
cat(unlist(chunks), sep = '\n')
```

```{r mean_coverage}
cov %>%
  dplyr::filter(lvl == 'drivers_samtools_depth') %>%
  dplyr::group_by(sample, gene, feature) %>%
  dplyr::summarise(mean_depth = mean(depth)) %>%
  ggplot(aes(x = sample, y = mean_depth, fill = feature)) +
  geom_col(position = 'dodge') +
  facet_grid(gene ~ .) +
  theme_classic()
```