### 2024-01-31 infercnv/reports/infercnv.Rmd
add infercnv rmd to show all plots, changes to driver coverage rmd

---
title: "Copy number variants in rheumatoid arthritis data from inferCNV"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: true
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
    rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') |> paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/infercnv/') ; setwd(wd)
# rmarkdown::render('reports/infercnv.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = T))

knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE,
  cache.path = 'reports/infercnv_cache/')

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)
library(slickR)

# dev.off() if there is a device
dev_off_if <- function() {if (length(dev.list()) != 0) { dev.off() }}
```

inferCNV was run on the scRNA-seq data to call copy number variants in synovial cells. No reference group was used, so the average profile of all cells was used in each patient as the baseline.

```{r, results = 'asis'}
pngs <-
  list.files('out/Zhang2023/', pattern = '^BRI') %>% head %>%
  purrr::set_names() %>%
  purrr::map(function(sample_id) {
    list.files(paste0('out/Zhang2023/', sample_id), pattern = '.png', 
                     recursive = T, full.names = T) %>%
      {setNames(., basename(.))}
  })

chunks <- c()

for (sample_id in names(pngs)) {

  chunks <- c(chunks, paste0('\n#### ', sample_id))

  # for (i in names(pngs[[sample_id]])) {
  #   
  #   chunks <- c(chunks, paste0('\n##### ', i))

    knit_plot <- c(
      '',
      '```{r, warning = F, message = F, echo = F}',
      paste0('slickR(obj = pngs[["', sample_id, '"]]) %synch% ',
             '( slickR(names(pngs[["', sample_id, '"]]), slideType = "p") + settings(arrows = F) )'),
      '```',
      ''
    )
    knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
    dev_off_if()

  }

  knit_plot <- c(
    '',
    '```{r, warning = F, message = F, echo = F}',
    paste0('slickR(obj = pngs[["', sample_id, '"]]) %synch% ',
           '( slickR(names(pngs[["', sample_id, '"]]), slideType = "p") + settings(arrows = F) )'),
    '```',
    ''
  )
  knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
  dev_off_if()

  chunks <- c(chunks, knitted_plot)

}

cat(unlist(chunks), sep = '\n')
```