### 2024-02-05 scomatic/reports/driver_coverage.Rmd

---
title: "Driver coverage"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: false
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
  id: NULL
  depths: NULL
  drivers: NULL
  ref_cds: NULL
  cache_dir: './driver_coverage_cache/'
  rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') |> paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/') ; setwd(wd)
# rmarkdown::render('reports/driver_coverage.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = F))

knitr::opts_knit$set(root.dir = './')
knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE,
  cache.path = params$cache_dir)

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)

# dev.off() if there is a device
dev_off_if <- function() {if (length(dev.list()) != 0) { dev.off() }}
```

We read in drivers and their depths.

```{r read_cov}
# read coverage
cov <- xfun::cache_rds({
  cov <- 
    readr::read_tsv('depths.tsv') %>% 
    dplyr::mutate(chr = as.character(chr))
  cov
}, file = 'cov.rds', rerun = params$rerun)

# read drivers
drivers <- readr::read_tsv(params$drivers)
```

### Annotate exons and introns

We use the reference object from DNDSCV to delimit intronic and exonic regions of the genes.

```{r features}
# get ref_cds object from DNDSCV
ref_cds <- 
  readRDS(params$ref_cds)

# annotate gene features
cov <- xfun::cache_rds({

  # get all positions in each gene
  gene_spans <-
    ref_cds[c(drivers$gene)] %>%
    purrr::map(function(i) {
      tibble::tibble(
        chr = as.character(i$chr),
        start = i$intervals_cds[, 1],
        end = i$intervals_cds[, 2]) %>%
        dplyr::mutate(exon = paste0('exon', dplyr::row_number()))
    })

  # drop those not found
  gene_spans[is.na(names(gene_spans))] <- NULL

  # annotate exons
  exons <-
    gene_spans %>%
    dplyr::bind_rows(.id = 'gene') %>%
    dplyr::mutate(pos = purrr::map2(start, end, ~ .x:.y)) %>%
    tidyr::unnest(cols = 'pos') %>%
    dplyr::select(-start, -end)
  cov <-
    cov %>%
    dplyr::left_join(exons, by = c('chr', 'pos', 'gene')) %>%
    dplyr::mutate(exon = ifelse(is.na(exon), 'intron', exon),
                  feature = ifelse(exon == 'intron', 'intron', 'exon'))

  # return
  cov

}, file = 'cov_w_features.rds', rerun = params$rerun)
```

```{r plots}
# get exon coords for plotting
exon_coords <- xfun::cache_rds({
  exon_coords <-
    cov %>%
    dplyr::filter(feature == 'exon') %>%
    dplyr::group_by(gene, exon) %>%
    dplyr::summarise(start = min(pos), 
                     end = max(pos))
  exon_coords
}, file = 'exon_coords.rds', rerun = params$rerun)

# function: plot all
plot_all <- function(p_df, p_exon_coords, p_title) {
  ggplot() +
    geom_rect(
      data = p_exon_coords,
      aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
      colour = 'lightblue', fill = 'lightblue'
    ) +
    geom_col(
      data = p_df,
      aes(x = pos, y = depth, fill = feature)) +
    theme_classic() +
    theme(panel.border = element_rect(fill = NA),
          panel.spacing = unit(.05, 'lines')) +
    scale_fill_manual(values = c(exon = 'darkblue', intron = 'darkgrey')) +
    labs(title = paste(p_title, '(all)')) +
    facet_grid(celltype ~ ., space = 'free_x', scales = 'free_x') 
}

# function: plot exons 
plot_exons <- function(p_df, p_title) {
  p_df %>%
    # get only exonic sequence
    dplyr::filter(feature == 'exon') %>%
    # fix exon ordering
    dplyr::mutate(
      exon = exon %>% forcats::fct_reorder(exon %>% gsub('exon', '', .) %>% as.numeric())
    ) %>%
    ggplot(aes(x = pos, y = depth, group = celltype)) +
    geom_col(fill = 'darkblue') +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.spacing = unit(.05, 'lines'),
          panel.border = element_rect(fill = NA)) +
    labs(title = paste(p_title, '(exons)')) +
    facet_grid(celltype ~ exon, space = 'free_x', scales = 'free_x') 
}

# create list of plots by gene x sample
p <- xfun::cache_rds({

  # initiate plot list
  p <- list()
  n_cts <- list()

  cov %>% 
    dplyr::group_by(gene) %>%
    dplyr::group_split() %>%
    purrr::map(function(df_i) {

      # for testing:
      # cov %>% dplyr::group_by(gene) %>% dplyr::group_split() %>% {.[[1]]} -> df_i

      # get info
      gene_i <- unique(df_i$gene)
      p[[gene_i]] <<- list()

      # get number of celltypes for plot dims
      p[[gene_i]]$n_cts <<- 
        df_i %>% 
        dplyr::filter(!is.na(celltype)) %>% 
        dplyr::pull(celltype) %>% 
        dplyr::n_distinct()

      # plot by sample and by celltype
      df_i %>%
        dplyr::mutate(suffix = dplyr::case_when(is.na(celltype) ~ '',
                                                TRUE ~ '_by_celltype')) %>%
        {split(., .$suffix)} %>%
        purrr::map(function(lvl) {

          # get suffix
          p_suffix <- unique(lvl$suffix)

          # plot all
          p[[gene_i]][[paste0('all', p_suffix)]] <<-
            plot_all(
              p_df = lvl,
              p_exon_coords = exon_coords %>% dplyr::filter(gene == gene_i),
              p_title = paste0(gene_i, p_suffix)
            )

          # plot exons
          p[[gene_i]][[paste0('exons', p_suffix)]] <<-
            plot_exons(
              p_df = lvl,
              p_title = paste0(gene_i, p_suffix)
            )
        })
    })

  # return 
  p

}, file = 'plots.rds', rerun = params$rerun)
```

```{r knit_plots, echo = F, warning = F, message = F, results = 'asis'}
chunks <- c()

for (info in names(p)) {
  chunks <- c(chunks, paste0('\n####', info))
  message(info)

  knit_plot <- c(
    '',
    paste0('```{r ', info, ', class.source = "fold-hide", warning = F, message = F, ',
           'fig.dim = c(20, ', 4 + 4 * p[[info]]$n_cts, ')}'),
    paste0('p_i <- p[["', info, '"]]'),
    paste0('grid.arrange(p_i$all, p_i$all_by_celltype, p_i$exons, p_i$exons_by_celltype, ncol = 1, ',
           'heights = c(2, p_i$n_cts * 2, 2, p_i$n_cts * 2))'),
    '```',
    '')
  knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
  dev_off_if()

  chunks <- c(chunks, knitted_plot)
}
dev_off_if()

# evaluate chunks
cat(unlist(chunks), sep = '\n')
```

```{r mean_coverage}
cov %>%
  dplyr::filter(lvl == 'drivers_samtools_depth') %>%
  dplyr::group_by(sample, gene, feature) %>%
  dplyr::summarise(mean_depth = mean(depth)) %>%
  ggplot(aes(x = sample, y = mean_depth, fill = feature)) +
  geom_col(position = 'dodge') +
  facet_grid(gene ~ .) +
  theme_classic()
```