### 2024-02-07 infercnv/reports/infercnv.Rmd
change knit infercnv back to manual integration of all pngs

---
title: "Copy number variants in rheumatoid arthritis data from inferCNV"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: false
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
  infercnv_dir: NULL
  rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') |> paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/infercnv/') ; setwd(wd)
# rmarkdown::render('reports/infercnv.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = T))

# chunk options
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE)

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)
library(slickR)

# dev.off() if there is a device
dev_off_if <- function() {if (length(dev.list()) != 0) { dev.off() }}
```

inferCNV was run on the scRNA-seq data to call copy number variants in synovial cells. No reference group was used, so the average profile of all cells was used in each patient as the baseline.

```{r, results = 'asis'}
pngs <-
  list.files(params$infercnv_dir,
             pattern = '^BRI', 
             include.dirs = T) %>%
  purrr::set_names() %>%
  purrr::map(function(id) {
    # collect pngs
    lis <-
      list.files(file.path(params$infercnv_dir, id), 
                 pattern = '.png', 
                 recursive = T, 
                 full.names = T) %>%
      {setNames(., gsub('infercnv.', '', tools::file_path_sans_ext(basename(.))))}

    # reorder (final plot first)
    c(lis['infercnv'], lis[names(lis) != 'infercnv'])
  })

chunks <- c()

for (id in names(pngs)) {

  chunks <- c(chunks, paste0('\n### ', id))

  for (i in names(pngs[[id]])) {

    chunks <- c(chunks, paste0('\n#### ', i))

    knit_plot <- c(
      '',
      '```{r, warning = F, message = F, echo = F}',
      paste0('knitr::include_graphics("../', pngs[[id]][[i]], '")'),
      '```',
      ''
    )
    knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
    dev_off_if()

    chunks <- c(chunks, knitted_plot)

  }
}

cat(unlist(chunks), sep = '\n')
```