### 2024-02-28 scomatic/reports/driver_coverage.Rmd

---
title: "Driver coverage"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: false
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
  drivers: NULL
  ref_cds: NULL
  cache_dir: './driver_coverage_cache/'
  rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') |> paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/') ; setwd(wd)
# rmarkdown::render('reports/driver_coverage.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = F))

knitr::opts_chunk$set(
  warning = FALSE, 
  dpi = 300, 
  message = FALSE,
  cache.path = params$cache_dir)

# libraries
library(magrittr)
library(ggplot2)
library(gridExtra)
library(GenomicRanges)
library(BSgenome)

# dev.off() if there is a device
dev_off_if <- function() {if (length(dev.list()) != 0) { dev.off() }}
```

We read in drivers and their coverage by sample / celltype.

```{r read_cov}
# read drivers
drivers <- readr::read_tsv(params$drivers)

# read coverage
cov <- xfun::cache_rds({

  cov <-
    list.files(pattern = '_coverage.tsv') %>% 
    purrr::set_names(., gsub('_coverage.tsv', '', .)) %>%
    purrr::map(function(file) {

      # read  
      df <- 
        readr::read_tsv(file, show_col_types = F) %>%
        dplyr::rename(all = `NA`)

      # reformat for GenVisR
      df %>%
        dplyr::select(-c(chr, pos, gene)) %>%
        colnames() %>%
        purrr::set_names(., .) %>%
        purrr::map(function(ct) {
          df %>%
            dplyr::transmute(
              gene,
              chromosome = paste0('chr', chr), 
              end = pos, 
              cov = get(ct))

          }) %>%
        dplyr::bind_rows(.id = 'celltype')

    }) %>%
    dplyr::bind_rows(.id = 'id') %>%
    split(., .$gene)

  # return
  cov

}, file = 'cov.rds', rerun = params$rerun)
```

We also load the `BSgenome` and `TxDb` objects.

```{r bsgenome}
# BSgenome
library("BSgenome.Hsapiens.UCSC.hg38")
genome <- BSgenome.Hsapiens.UCSC.hg38

# TxDb
library("TxDb.Hsapiens.UCSC.hg38.knownGene")
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

Then we plot the coverage.

```{r gencov}
plot_cov <- function(cov, gene, var, expand_features_of_interest = T) {

  # split by variable(s)
  x <-
    cov[[gene]] %>%
    dplyr::select(chromosome, end, cov) %>%
    split(., cov[[gene]][, var])

  # make granges obj
  gr <-
    GRanges(
      seqnames = as.character(unique(cov[[gene]]$chromosome)),
      ranges = IRanges(start = min(cov[[gene]]$end) - 500,
                       end = max(cov[[gene]]$end) + 500))

  # check what to transform
  transform <- c('Intron', 'CDS', 'UTR')
  base <- c(10, 2, 2)
  if (expand_features_of_interest == T) { base <- transform <- NA }

  # plot
  GenVisR::genCov(
    x = x,
    txdb = txdb, 
    gr = gr, 
    genome = genome,
    cov_plotType = 'line',
    cov_plotLayer = ggplot2::theme(axis.title.y = ggplot2::element_blank()),
    label_txtSize = 4,
    lab2plot_ratio = c(2, length(x)),
    gene_labelTranscriptSize = 2,
    gene_name = gene,
    base = base,
    transform = transform)

}

plot_cov(cov, 'ARID2', 'celltype') 
plot_cov(cov, 'ARID2', 'celltype', F)
plot_cov(cov, 'ARID2', 'id')
plot_cov(cov, 'ARID2', 'id', F)
```

```{r exit}
knitr::knit_exit()
```

### Annotate exons and introns

We use the reference object from DNDSCV to delimit intronic and exonic regions of the genes.

```{r features}
# get ref_cds object from DNDSCV
ref_cds <- 
  readRDS(params$ref_cds)

# annotate gene features
cov <- xfun::cache_rds({

  # only drivers in ref_cds
  avail_drivers <-
    drivers %>%
    dplyr::filter(gene %in% names(ref_cds))

  # get all positions in each gene
  gene_spans <-
    ref_cds[c(avail_drivers$gene)] %>%
    purrr::map(function(i) {
      tibble::tibble(
        chr = as.character(i$chr),
        start = i$intervals_cds[, 1],
        end = i$intervals_cds[, 2]) %>%
        dplyr::mutate(exon = paste0('exon', dplyr::row_number()))
    })

  # annotate exons
  exons <-
    gene_spans %>%
    dplyr::bind_rows(.id = 'gene') %>%
    dplyr::mutate(pos = purrr::map2(start, end, ~ .x:.y)) %>%
    tidyr::unnest(cols = 'pos') %>%
    dplyr::select(-start, -end)
  cov <-
    cov %>%
    dplyr::left_join(exons, by = c('chr', 'pos', 'gene')) %>%
    dplyr::mutate(exon = ifelse(is.na(exon), 'intron', exon),
                  feature = ifelse(exon == 'intron', 'intron', 'exon'))

  # return
  cov

}, file = 'cov_w_features.rds', rerun = params$rerun)
```

```{r plots}
# get exon coords for plotting
exon_coords <- xfun::cache_rds({
  exon_coords <-
    cov %>%
    dplyr::filter(feature == 'exon') %>%
    dplyr::group_by(gene, exon) %>%
    dplyr::summarise(start = min(pos), 
                     end = max(pos))
  exon_coords
}, file = 'exon_coords.rds', rerun = params$rerun)

# function: plot all
plot_all <- function(p_df, p_exon_coords, p_title) {
  ggplot() +
    geom_rect(
      data = p_exon_coords,
      aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
      colour = 'lightblue', fill = 'lightblue'
    ) +
    geom_col(
      data = p_df,
      aes(x = pos, y = depth, fill = feature)) +
    theme_classic() +
    theme(panel.border = element_rect(fill = NA),
          panel.spacing = unit(.05, 'lines')) +
    scale_fill_manual(values = c(exon = 'darkblue', intron = 'darkgrey')) +
    labs(title = paste(p_title, '(all)')) +
    facet_grid(celltype ~ ., space = 'free_x', scales = 'free_x') 
}

# function: plot exons 
plot_exons <- function(p_df, p_title) {
  p_df %>%
    # get only exonic sequence
    dplyr::filter(feature == 'exon') %>%
    # fix exon ordering
    dplyr::mutate(
      exon = exon %>% forcats::fct_reorder(exon %>% gsub('exon', '', .) %>% as.numeric())
    ) %>%
    ggplot(aes(x = pos, y = depth, group = celltype)) +
    geom_col(fill = 'darkblue') +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.spacing = unit(.05, 'lines'),
          panel.border = element_rect(fill = NA)) +
    labs(title = paste(p_title, '(exons)')) +
    facet_grid(celltype ~ exon, space = 'free_x', scales = 'free_x') 
}

# function: plot stacked exon coverage
plot_exons_stacked <- function(p_df, p_title) {
  p_df %>% 
    # get only exonic sequence
    dplyr::filter(feature == 'exon') %>%
    # fix exon ordering
    dplyr::mutate(
      exon = exon %>% gsub('exon', '', .) %>% as.numeric(),
      facets = ifelse(is.na(celltype), 'all', 'by celltype')) %>%
    ggplot(aes(x = pos, y = depth, group = celltype, fill = celltype)) +
    geom_col(position = 'stack') +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.spacing = unit(.05, 'lines'),
          panel.border = element_rect(fill = NA)) +
    labs(title = paste(p_title, '(exons)')) +
    facet_grid(facets ~ exon, space = 'free_x', scales = 'free_x') 
}

# create list of plots by gene x sample
p <- xfun::cache_rds({

  # initiate plot list
  p <- list()
  n_cts <- list()

  cov %>% 
    {split(., .$id)} %>%
    purrr::map(function(df_id) {

      # get id, inititate sublist
      id_i <- unique(df_id$id)
      p[[id_i]] <<- list()

      # go by gene
      df_id %>%
        {split(., .$gene)} %>%
        purrr::map(function(df_i) {

          # get gene, initiate sublist
          gene_i <- unique(df_gene$gene)
          p[[id_i]][[gene_i]] <<- list()

          # get number of celltypes for plot dims
          p[[id_i]][[gene_i]]$n_cts <<- 
            df_i %>% 
            dplyr::filter(!is.na(celltype)) %>% 
            dplyr::pull(celltype) %>% 
            dplyr::n_distinct()

          # plot by sample and by celltype
          df_i %>%
            dplyr::mutate(suffix = dplyr::case_when(is.na(celltype) ~ '',
                                                    TRUE ~ '_by_celltype')) %>%
            {split(., .$suffix)} %>%
            purrr::map(function(lvl) {

              # get suffix
              p_suffix <- unique(lvl$suffix)

              # plot all
              p[[id_i]][[gene_i]][[paste0('all', p_suffix)]] <<-
                plot_all(
                  p_df = lvl,
                  p_exon_coords = exon_coords %>% dplyr::filter(gene == i),
                  p_title = paste0(i, p_suffix)
                )

              # plot exons
              p[[id_i]][[gene_i]][[paste0('exons', p_suffix)]] <<-
                plot_exons(
                  p_df = lvl,
                  p_title = paste0(i, p_suffix)
                )

              # plot exons stacked
              p[[id_i]][[gene_i]][[paste0('exons_stacked', p_suffix)]] <<-
                plot_exons_stacked(
                  p_df = lvl,
                  p_title = paste0(i, p_suffix)
                )
            })
        })
    })

  # return 
  p

}, file = 'plots.rds', rerun = params$rerun)
```

```{r knit_plots, echo = F, message = F, results = 'asis', warning = T, error = T}
chunks <- c()

for (info in names(p)) {
  chunks <- c(chunks, paste0('\n#### ', info))
  message(info)

  knit_plot <- c(
    '',
    paste0('```{r ', info, ', class.source = "fold-hide", warning = F, message = F, ',
           'fig.dim = c(20, ', 4 + 4 * p[[info]]$n_cts, ')}'),
    paste0('p_i <- p[["', info, '"]]'),
    paste0('grid.arrange(p_i$all, p_i$all_by_celltype, p_i$exons, p_i$exons_by_celltype, ncol = 1, ',
           'heights = c(2, p_i$n_cts * 2, 2, p_i$n_cts * 2))'),
    '```',
    '')
  knitted_plot <- knitr::knit_child(text = knit_plot, envir = environment(), quiet = T)
  dev_off_if()

  chunks <- c(chunks, knitted_plot)
}
dev_off_if()

# evaluate chunks
cat(unlist(chunks), sep = '\n')
```

```{r mean_coverage, eval = F}
cov %>%
  dplyr::filter(lvl == 'drivers_samtools_depth') %>%
  dplyr::group_by(sample, gene, feature) %>%
  dplyr::summarise(mean_depth = mean(depth)) %>%
  ggplot(aes(x = sample, y = mean_depth, fill = feature)) +
  geom_col(position = 'dodge') +
  facet_grid(gene ~ .) +
  theme_classic()
```