### 2024-08-07 infercnv/src/nb15_nanoseqLollipopGenePlots_targ_nano_grch37_031023.Rmd

---
title: "nanoseqLollipopGenePlots"
author: "Matt Neville"
date: '2022-07-12'
output: html_document
---

# 1. Prep
## Paths/Libraries
```{r setup}
#Input files on the farm
dndsPath <- "~/Mounts/peter/liver/analysis/targeted_liver_panel/"
refCDSpath <- "~/Mounts/peter/liver/code/dNdScv/RefCDS_GRCh37_vF.Rdat"
lof_gof_path <- "~/Mounts/peter/liver/data/NanoSeq/lof_gof38.tsv"
nanoseqVarsPath <- "~/Mounts/peter/liver//analysis/targeted_liver_panel/combined_muts_annot.tsv"

#Local output
# outputPath <- "/Volumes/GoogleDrive/My Drive/Sperm/lolliplot/"
outputPath <- "/Users/nb15/Mounts/peter/liver/analysis/targeted_liver_panel/lolliplots/coverage/"

library(Biostrings)
library(GenomicRanges)
library(MetBrewer)
library(egg)
library(plyr)
library(tidyverse)
```

## Load Datasets
```{r loadDatasets}

## load dnds hg37 refCDS object from dNdS -> to get coding sequence intervals
load(refCDSpath)

# Katrina's clinvar annotated LOF/GOF list
lof_gof <- read_tsv(lof_gof_path, col_types = cols()) %>%
  dplyr::rename(mut = alt)

## dataframe that holds list number in RefCDS object for each gene; so you can easily extract your gene of interest cds intervals
genes <- as.data.frame(matrix(nrow=length(RefCDS), ncol=2))
colnames(genes) <- c("gene_name","row_in_refcds")
for( i in 1:length(RefCDS)){
  genes$gene_name[i] <- RefCDS[[i]]$gene_name
  genes$row_in_refcds[i] <- i
}

## load dnds object to annotate muts and colour by impact
annotmuts <- readRDS(paste0(dndsPath, "dnds_all_dc.rds"))$annotmuts

## merge nanoseq_data with dnds annotations and clinvar annotations
nanoseq_data <- read_tsv(nanoseqVarsPath, col_types = cols()) %>% 
  dplyr::select(sampleID, chr, pos, ref, mut, duplex_cov) %>% 
  mutate(sampleID = as.character(sampleID)) %>%
  mutate(chr = gsub("chr","",chr)) %>%
  left_join(annotmuts,by=c("sampleID","chr","pos","ref","mut")) %>%  
  left_join(lof_gof,by=c("chr","pos","ref","mut")) %>%  
  mutate(label = if_else(fxn == "altered_function", "G", "")) %>%
  mutate(label = if_else(fxn == "LOF", paste0(label,"L"), label)) %>%
  # rename cols to match nanoseq_data colnames
  dplyr::rename(CHROM = chr,
         POS = pos,
         REF = ref,
         ALT = mut) %>% 
  #Change label based on clinvar Katrina annotation
  # mutate(label = if_else(fxn == "altered_function", "G", "")) %>% 
  # mutate(label = if_else(fxn == "LOF", paste0(label,"L"), label)) %>% 
  mutate(impact = replace_na(impact, "Non-coding")) %>% 
  mutate(impact = str_replace(impact, "no-SNV", "Indel"))  %>% 
  mutate(name = if_else(impact == "Indel", 
                        #Formating for indels
                        replace_na(str_remove(ntchange, ".*-"), ".") %>% 
                          str_replace("frshift", "FS") %>% 
                          str_replace("inframe", "IF"), 
                        #Name for SNVs
                        replace_na(aachange, ".")))

## define genes of interest by dNdS sig genes or by custom list or put in gene names (example below)
#goi <- readRDS(paste0(dndsPath, "spermTargdNdS.RDS"))$sel_cv %>% 
  # dplyr::filter(qglobal_cv < 0.05 | qallsubs_cv < 0.05) %>% 
#  pull(gene_name)

#goi <- c("FOXO1","KLF15","CYP2E1","ACVR2A")
goi <- c("A1CF","TNRC6B","FOXO1","GPAM","ACVR2A","TF","CHD4")

```

# 2. Coverage
## Calculate cov
Bash code on farm to generate the total coverage across all samples in cohort
```{bash burdens, eval = F}

#Go to targeted directory, create folders
cd /lustre/scratch126/casm/team154pc/nb15/liver/data/NanoSeq/plate_9_DN942040D_exome_1/Analysis
mkdir -p cov cov/cut cov/source cov/jobLogs

#Unzip  and rename all nanoseq coverage files to single folder then
#Format files to remove trinuc context
for sample in `ls variants_subs`;
  do echo $sample
  cp variants_subs/$sample/results.cov.bed cov/source/$sample.bed
  sed 's/;/\t/g' cov/source/$sample.bed | cut -f 1-3,6 > cov/cut/$sample.bed
done

#Make summary files with coverage for all indivs at all positions
cd /lustre/scratch126/casm/team154pc/nb15/liver/data/NanoSeq/plate_9_DN942040D_exome_1/Analysis/cov/cut
bsub -G team154-grp -o ../jobLogs/sumCov.%J.out -e ../jobLogs/sumCov.%J.err -q normal -R 'select[mem>=1000] rusage[mem=1000] span[hosts=1]' -M 1000 'module load bedtools; bedtools unionbedg -i $(ls *.bed) -header -names $(ls *.bed) > ../summaryCov.bed'

#Get sum of coverage at each site
cd /lustre/scratch126/casm/team154pc/nb15/liver/data/NanoSeq/plate_9_DN942040D_exome_1/Analysis/cov

#These 2 lines need to be in a bash script so could for instance type 'nano getTotalCov.sh', paste these 2 lines in and then make it executable with 'chmod +x getTotalCov.sh'

#!/bin/bash
awk 'OFS="\t" {sum=0; for (col=4; col<=NF; col++) sum += $col; print $1, $2, $3, sum;}' summaryCov.bed | tail -n +2 > totalCov.bed

bsub -G team154-grp -q normal -o jobLogs/sumCov.%J.out -e jobLogs/sumCov.%J.err -R 'select[mem>=1000] rusage[mem=1000] span[hosts=1]' -M 1000 ./getTotalCov.sh

```

## Read into R
```{r cov}
#Coverage summaries
totTargCov <- read_tsv("~/Mounts/peter/liver/data/NanoSeq/plate_9_DN942040D_exome_1/Analysis/cov/totalCov.bed", col_types = "c-dd", col_names = c("chr", "pos", "cov"))

```

# 3. Plot
## Plotting functions
```{r cov}
# jitterLables and reAdjustLabels adapted from lolliplots package
jitterLables <- function(coor, xscale, lineW, weight=1.2){
    if(weight==1.2) {
      stopifnot("Please sort your inputs by start position"= 
                  order(coor)==1:length(coor))
    }
    if(weight<0.5) return(coor)
    stopifnot(length(xscale)==2)
    pos <- (coor-1)/(xscale[2]-1)
    # pos <- convertX(unit(coor01, "native"), "npc", valueOnly=TRUE)
    pos.diff <- diff(c(0, pos, 1))
    idx <- which(pos.diff < weight*lineW)
    if(length(idx)<1){
        return(coor)
    }
    if(all(idx %in% c(1, length(pos)+1))){
        return(coor)
    }
    idx.diff <- diff(c(-1, idx))
    idx.grp <- rle(idx.diff)
    idx.grp$values[idx.grp$values==1] <- length(pos) + 1:sum(idx.grp$values==1)
    idx.grp <- inverse.rle(idx.grp)
    idx.grp.w <- which(idx.grp>length(pos))-1
    idx.grp.w <- idx.grp.w[idx.grp.w>0]
    idx.grp[idx.grp.w] <- idx.grp[idx.grp.w+1]
    idx.grp <- split(idx, idx.grp)
    flag <- as.numeric(names(idx.grp))>length(pos)
    idx.grp.mul <- lapply(idx.grp[flag], function(.ele){
        c(.ele[1]-1, .ele)
    })
    idx.grp.sin <- lapply(idx.grp[!flag], function(.ele){
        lapply(as.list(.ele), function(.ele){c(.ele-1, .ele)})
    })
    idx.grp.sin <- unlist(idx.grp.sin, recursive = FALSE)
    idx.grp <- c(idx.grp.mul, idx.grp.sin)

    adj.pos <- lapply(idx.grp, function(.ele){
        .ele <- .ele[.ele>0 & .ele<=length(pos)]
        this.pos <- pos[.ele]
        names(this.pos) <- .ele
        if(length(this.pos)%%2==1){
            center <- ceiling(length(this.pos)/2)
        }else{
            center <- length(this.pos)/2 + .5
        }
        if(length(this.pos)>5){ ## too much, how to jitter?
            this.pos <- this.pos + 
                ((1:length(this.pos))-center) * (weight-.1) * 
                lineW/ceiling(log(length(this.pos), 5))
        }else{
            this.pos <- this.pos + 
                ((1:length(this.pos))-center) * (weight-.1) * lineW
        }
        this.pos
    })
    names(adj.pos) <- NULL
    adj.pos <- unlist(adj.pos)
    coor[as.numeric(names(adj.pos))] <- adj.pos*diff(xscale)+xscale[1]

    Recall(coor, xscale=xscale, lineW=lineW, weight=weight-0.2)
}

reAdjustLabels <- function(coor, lineW, xmax){
  # resort
  coor <- sort(coor)
  bins <- ceiling(1/lineW)
  pos <- (coor-1)/(xmax-1)
  # pos <- convertX(unit(coor, "native"), "npc", valueOnly=TRUE)
  pos.bin <- cut(pos, c(-Inf, (0:bins)*lineW, Inf), labels=0:(bins+1), right=FALSE)

  ## split the coors by into clusters
  ## give the clusters with more idx more spaces if there are spaces between clusters
  tbl <- table(pos.bin)
  if(all(tbl<2)) return(coor)
  tbl.len <- length(tbl)
  if(tbl.len<3) return(coor)
  loops <- 1000
  loop <- 1
  while(any(tbl==0) && any(tbl>1) && loop < loops){
    tbl.bk <- tbl
    for(i in order(tbl.bk, decreasing=TRUE)){
      if(tbl[i]>1 && tbl.bk[i]==tbl[i]){
        if(i==1){
          if(tbl[2]<tbl[1]){
            half <- sum(tbl[1:2])/2
            tbl[2] <- ceiling(half)
            tbl[1] <- floor(half)
          }
        }else{
          if(i==tbl.len){
            if(tbl[tbl.len]>tbl[tbl.len-1]){
              half <- sum(tbl[(tbl.len-1):tbl.len])/2
              tbl[tbl.len-1] <- ceiling(half)
              tbl[tbl.len] <- floor(half)
            }
          }else{
            if(tbl[i-1]<tbl[i+1]){
              ## i-1 and i should be balanced
              half <- sum(tbl[(i-1):i])/2
              tbl[i-1] <- floor(half)
              tbl[i] <- ceiling(half)
            }else{
              half <- sum(tbl[i:(i+1)])/2
              tbl[i] <- floor(half)
              tbl[i+1] <- ceiling(half)
            }
          }
        }
      }
    }
    loop <- loop + 1
  }
  coef <- unlist(lapply(tbl, function(.ele){
    if(.ele==0) return(0)
    .ele <- seq(from=0, to=1, length.out=.ele+1)
    (.ele[-length(.ele)] + .ele[-1])/2
  }))
  coef <- coef[coef!=0]
  coor <- (rep(as.numeric(names(tbl)), tbl) - 1 + coef) * lineW
  # coor <- convertX(unit(coor, "npc"), "native", valueOnly=TRUE)
  coor = ((xmax-1) * coor) + 1
  coor
}

# Covert genomic bp to coding sequence coordinate
getCDSindex <- function(intervals) {
  strand <- intervals$strand[1]
  allBP <- tibble()
  for(exon in 1:length(intervals$length)) {
    exonBP <- tibble(pos = seq(intervals$start[exon], intervals$end[exon]))
    allBP <- bind_rows(allBP, exonBP) 
  }
  # Forward strand standard
  if(strand == 1) {
  allBP <- allBP %>% 
      mutate(cdsPos = seq(1:sum(intervals$length)))    
  } else {
    #Reverse strand flip cds coords so all genes show left to right
    allBP <- allBP %>% 
      mutate(cdsPos = seq(from = sum(intervals$length), to = 1))  
  }
  return(allBP)
}

#For a mutation outside the CDS (splice or non-coding) set its CDS position as .5 off the nearest exon border
getCDSindexNonCDS <- function(cdsCoordsIndex, vars, strand) {
  nonCDSvars <- vars %>% filter(impact %in% c("Essential_Splice", "Non-coding") | (impact == "Indel" & is.na(ntchange)))
  closestCDS <- cdsCoordsIndex %>% dplyr::slice(rep(1:n(), each = dim(nonCDSvars)[1])) %>% 
    mutate(nonCDSpos = rep(nonCDSvars$POS, times = dim(cdsCoordsIndex)[1])) %>% 
    mutate(gap = nonCDSpos - pos) %>% 
    mutate(absGap = abs(gap)) %>% 
    group_by(nonCDSpos) %>% filter(absGap == min(absGap)) %>% 
    ungroup() %>% 
    #flip direction if reverse strand
    mutate(gap = gap * strand) %>% 
    #Set offset based on strand direction
    mutate(cdsPos = if_else(gap < 0, cdsPos - 0.5, cdsPos + 0.5)) %>% 
    select(POS = nonCDSpos, cdsPos)
  return(closestCDS)
}
```

## Lolliplot
Loops through selected genes, outputs file
```{r lolliplot}

#Plot Settings
#Set height of coverage y-axis
yMax = max(totTargCov$cov)+500

#True means drop non-coding (intronic) mutations, False means keep (will show at exon borders)
filterNonCoding = T
#Controls how spaced variant points will be. If you change pdf output size, you may need to adjust this
pointWidth = 0.017

## loop over genes
for(g in 1:length(goi)){
g_CDS <- RefCDS[[which(genes$gene_name==goi[g])]]

intervals_cds0 <- tibble(chr = g_CDS$chr,
                        strand = g_CDS$strand,
                        start = g_CDS$intervals_cds[,1],
                        end = g_CDS$intervals_cds[,2]) %>% 
  arrange(start) %>% 
  mutate(length = (end + 1) - start)

cdsCoordsIndex <- getCDSindex(intervals_cds0)

intervals_cds <- intervals_cds0 %>% 
  left_join(cdsCoordsIndex %>% select(start = pos, cdsStart = cdsPos), by = "start") %>% 
  left_join(cdsCoordsIndex %>% select(end = pos, cdsEnd = cdsPos), by = "end")

# Coverage
geneCov <- cdsCoordsIndex %>% 
  left_join(totTargCov %>% select(-chr), by = "pos") %>% 
  mutate(cov = replace_na(cov, 0))

# subset nanoseq_data to variants in the gene
# Can choose to filter non-coding variants or not
gene0 <- nanoseq_data %>% filter(CHROM == intervals_cds$chr[1] & POS >= min(intervals_cds %>% pull(start)) & POS <= max(intervals_cds %>% pull(end)))

gene <- gene0 %>% 
  left_join((cdsCoordsIndex %>% dplyr::rename(POS = pos) %>% bind_rows(getCDSindexNonCDS(cdsCoordsIndex, gene0, strand = intervals_cds$strand[1]))), by = "POS") %>% 
  arrange(cdsPos) %>% 
  drop_na(cdsPos)

  if(dim(gene)[1] == 0) {
    xMax <- max(c(intervals_cds$cdsEnd,intervals_cds$cdsStart)) + max(c(intervals_cds$cdsEnd,intervals_cds$cdsStart))*0.05
    xMin <- 1 - xMax*0.01
    pCov <- ggplot(geneCov, aes(x = cdsPos, y = cov)) + 
      #Coverage area
      geom_area(fill = '#ccece6') + ylab('Duplex_Cov') + 
      scale_x_continuous(expand = c(0, 0), limits = c(xMin, xMax)) +
      coord_cartesian(clip = "off") +
      scale_y_continuous(limits = c(0, yMax), expand = c(0, 0)) +
      ggtitle(goi[g]) +
      theme(
        panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"),
        panel.spacing = unit(c(0, 0, 0, 0), "cm"),
        axis.text.y = element_text(margin = margin(0,0.3,0,0.2, "cm")),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.ticks.length.x = unit(0, "null"),
        legend.position = c(0.5, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.key = element_rect(fill = NA)
      )

    pCDS <- ggplot(intervals_cds, aes(xmin = cdsStart, xmax = cdsEnd, ymin = 0, ymax = 1)) +
      geom_rect(fill = "#d0d1e6", colour = "black") +
      theme_void() + 
      xlab("CDS") +
      scale_x_continuous(expand = c(0, 0), limits = c(xMin,xMax)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme(plot.margin = unit(c(0, 0.5, 0.5, 0.5), "cm"))

    ggarrange(pCov, pCDS, heights = c(15,1), nrow = 2, ncol = 1)

    filename <- paste0(outputPath, goi[g],"_lollipop.pdf")
    pdf(file = filename, height = 3, width = 11, onefile=FALSE)
    ggarrange(pCov, pCDS, heights = c(15,1), nrow = 2, ncol = 1)
    dev.off()
  }else {
  #Set colours for variant csq
  cols <- met.brewer("Archambault",5)
  gene$col <- NA
  gene$col[which(gene$impact=="Non-coding")] <- "gray100"
  gene$col[which(gene$impact=="Synonymous")] <- "gray76"
  gene$col[which(gene$impact=="Missense")] <- "#abd9e9"
  gene$col[which(gene$impact=="Nonsense")] <- "#fdae61"
  gene$col[which(gene$impact=="Indel")] <- cols[3]
  gene$col[which(gene$impact=="Essential_Splice")] <- "#fee090"
  gene$col[which(gene$impact=="Stop_loss")] <- cols[5]

  # Set label position with lolliplot functions
  lab.pos <- jitterLables(gene %>% arrange(cdsPos) %>% pull(cdsPos), 
                          xscale=c(1, max(cdsCoordsIndex %>% pull(cdsPos))), 
                          lineW=pointWidth)

  lab.pos <- reAdjustLabels(lab.pos, lineW=pointWidth, xmax = max(cdsCoordsIndex %>% pull(cdsPos)))

  variants <- gene %>% 
    mutate(labPos = lab.pos) %>% mutate(aachange = replace_na(aachange, ".")) %>%
    mutate(label = replace_na(label, " "))

  xMax <- max(c(intervals_cds$cdsEnd,intervals_cds$cdsStart)) + max(c(intervals_cds$cdsEnd,intervals_cds$cdsStart))*0.05

  xMin <- 1 - xMax*0.01

  pCov <- ggplot(geneCov, aes(x = cdsPos, y = cov)) + 
    #Coverage area
    geom_area(fill = '#ccece6') + ylab('Duplex_Cov') + 
    #Variant stems
    geom_segment(data = variants, aes(x=cdsPos, xend=cdsPos, y=0, yend= (0.125 * yMax))) +
    geom_segment(data = variants, aes(x=cdsPos, xend=labPos, y=(0.125 * yMax), yend=(0.25 * yMax))) +
    geom_segment(data = variants, aes(x=labPos, xend=labPos, y=(0.25 * yMax), yend=(0.325 * yMax))) +
    #Variant points
    geom_point(data = variants, aes(x = labPos, y = (0.325 * yMax), fill=col), shape=21,  color="black", size=5) +
    #Label aa change
    geom_text(data = variants, aes(label = name, x = labPos, y = (0.5 * yMax)), angle = 90) + 
    #Label clinvar
    geom_text(data = variants, aes(label = label, x = labPos, y = (0.325 * yMax)), size=3.5) +
    scale_fill_identity(guide = "legend", breaks = variants$col, labels = variants$impact) +
    scale_x_continuous(expand = c(0, 0), limits = c(xMin, xMax)) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(limits = c(0, yMax), expand = c(0, 0)) +
    ggtitle(goi[g]) +
    theme(
      panel.background = element_rect(fill = "transparent", colour = NA),
      plot.background = element_rect(fill = "transparent", colour = NA),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"),
      panel.spacing = unit(c(0, 0, 0, 0), "cm"),
      axis.text.y = element_text(margin = margin(0,0.3,0,0.2, "cm")),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.line = element_blank(),
      plot.title = element_text(hjust = 0.5),
      axis.ticks.length.x = unit(0, "null"),
      legend.position = c(0.5, 0.85),
      legend.direction = "horizontal",
      legend.title = element_blank(),
      legend.key = element_rect(fill = NA)
    )

  pCDS <- ggplot(intervals_cds, aes(xmin = cdsStart, xmax = cdsEnd, ymin = 0, ymax = 1)) +
    geom_rect(fill = "#d0d1e6", colour = "black") +
    theme_void() + 
    xlab("CDS") +
    scale_x_continuous(expand = c(0, 0), limits = c(xMin,xMax)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme(plot.margin = unit(c(0, 0.5, 0.5, 0.5), "cm"))

  ggarrange(pCov, pCDS, heights = c(15,1), nrow = 2, ncol = 1)

  filename <- paste0(outputPath, goi[g],"_lollipop.pdf")
  pdf(file = filename, height = 3, width = 11, onefile=FALSE)

  ggarrange(pCov, pCDS, heights = c(15,1), nrow = 2, ncol = 1)

  dev.off()
  }
}

```